---
title: "Posterior consistency in SLGP estimation: sampling approximations influence"
author: "Athénaïs Gautier"
date: "Last revised - October 2024"
output:
  html_document:
  #bookdown::pdf_document2 :
  #  keep_tex: FALSE
  #  toc: FALSE
  #  number_sections: true
    citation_package: natbib
#bibliography: bibliography.bib
header-includes:
 \usepackage{float}
 \usepackage{tikz}
 \usepackage{xcolor}
 \usepackage{amsmath, amsfonts, amsthm}
 \floatplacement{figure}{H}
 \usepackage{fvextra}
 \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}

---

\newcommand{\xX}{\mathbf{x}}
\newcommand{\xI}{\mathcal{T}}
\newcommand{\xR}{\mathbb{R}}
\newcommand{\dimI}{d_\xI}
\newcommand{\dimD}{d_D}
\newcommand{\dxx}{d_{\xX, \xX'}}
\newcommand{\kincrement}{k_{\text{inc}}}
\newcommand{\mincrement}{m_{\text{inc}}}
\newcommand{\xM}{M(\xX, \xX')}
\newcommand{\diam}{D_{\xX, \xX'}(\xI)}
\newcommand{\xY}{\Vert \xX - \xX' \Vert^{\alpha_1 /2}}
\newcommand{\sigFieldM}{\mathcal{B}(\xI)}

\newcommand{\proc}[3]{
\ifstrempty{#3}%
{%
\ifstrempty{#2}%
{%
#1
}{%
#1_{#2}
}%
}{%
(#1_{#2})_{#3}
}%
}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(viridis)
```


# Introduction

This document is about exploring the impact of our approximate sampling strategy for posterior consistency of Spatial Logistic Gaussian Processes (SLGPs) in estimating response distributions given predictor values. 

It comes as a complement to the main document titled "Posterior consistency in SLGP estimation".

```{r modelsload, include=FALSE}
library(SLGP)
load(file="SLGPsRef.Rdata")
```

We display the corresponding probability density functions (PDFs) for a few instances of the underlying indexing variable $x$. 

```{r reffigure, fig.cap = "Probability density fields (displayed for a few indices x) with different spatial regularity", fig.fullwidth=TRUE, fig.height=5, fig.width=10, fig.align='center' ,fig.pos="H", warning=FALSE, message=FALSE, echo=FALSE}

set.seed(123)
epsilon <- rnorm(SLGP1@p, sd=sqrt(SLGP1@hyperparams$sigma2))
SLGP1@coefficients <-
  SLGP2@coefficients <-
  SLGP3@coefficients <-
  SLGP4@coefficients <- as.matrix(t(epsilon))

dfGrid2 <- data.frame(expand.grid(seq(0, 1,, 1001), seq(0, 1,, 101)))
colnames(dfGrid2) <- c("t", "x")

pred1 <- predictSLGP_newNode(SLGPmodel = SLGP1,
                             newNodes = dfGrid2)
pred1$kernel <- "Lowest"

pred2 <- predictSLGP_newNode(SLGPmodel = SLGP2,
                             newNodes = dfGrid2)
pred2$kernel <- "Low"

pred3 <- predictSLGP_newNode(SLGPmodel = SLGP3,
                             newNodes = dfGrid2)
pred3$kernel <- "High"

pred4 <- predictSLGP_newNode(SLGPmodel = SLGP4,
                             newNodes = dfGrid2)
pred4$kernel <- "Highest"

scale_factor <- 0.05
dfplot <- rbind(pred1, pred2, pred3, pred4)
dfplot$kernel <- factor(paste0("Spatial regularity:\n", dfplot$kernel),
                        levels=paste0("Spatial regularity:\n", 
                                      c("Lowest", "Low", "High", "Highest")))

dfplot %>%
  dplyr::filter(x %in% seq(0, 1,, 26))%>%
  ggplot(aes(x=t, ymax=scale_factor*pdf_1+x, ymin=x, group=-x, fill=x))+
  geom_ribbon(col="grey", alpha=0.9)+  # Add grey borders around each tile
  scale_fill_viridis(option = "plasma",
                     guide = guide_colorbar(nrow = 1,
                                            title = "Indexing variable x",
                                            barheight = unit(2, units = "mm"),
                                            barwidth = unit(55, units = "mm"),
                                            title.position = 'top',
                                            label.position = "bottom",
                                            title.hjust = 0.5))+
  facet_grid(.~kernel)+
  theme_bw()+
  theme(legend.position = "bottom", axis.text.y = element_blank(),
        axis.ticks.y = element_blank(), axis.line.y = element_blank())+
  ylab("Rescaled probability density")+
  xlab("Response variable t")
```

We draw samples from each of these four reference probability density fields at various values of the indexing variable $x$. These samples will be used to perform the inference, enabling us to assess how well the SLGP-based models can recover the underlying distributions. 

We briefly investigate the 

```{r pardefine, include=FALSE}
nmax <- 1e5
nRep <- 5
```

# Generating samples with various schemes for the 
## Case 1: sampling x that are not truly uniform but on a coarse grid

```{r samplegeneration1, eval=FALSE}
nmax <- 1e5
nRep <- 1
i<- 1
for(i in seq(nRep)){
  for(nDiscret in c(51, 501)){
    for(nmax in round(10**seq(1, 5, 0.5))){
      
      title_file <- paste0("./samplingstudy/SLGPsSamp", i, 
                           "discret", nDiscret, "n", nmax,
                           "case1.Rdata")
      if(!file.exists(title_file)){
        deb<- Sys.time()
        cat(i, ".1\n")
        set.seed(i)
        x_candidates <- sample(seq(0, 1,, 101), nmax, replace=TRUE)
        set.seed(i)
        t <- as.data.frame(table(x_candidates))
        sampTemp <- sampleSLGP(SLGPmodel=SLGP1, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp<- data.frame(x=sampTemp$x,
                          t1 = sampTemp$t)
        set.seed(i)
        cat(i, ".2\n")
        sampTemp <- sampleSLGP(SLGPmodel=SLGP2, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t2<- sampTemp$t
        set.seed(i)
        cat(i, ".3\n")
        
        sampTemp <- sampleSLGP(SLGPmodel=SLGP3, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t3<- sampTemp$t
        set.seed(i)
        cat(i, ".4\n")
        
        sampTemp <- sampleSLGP(SLGPmodel=SLGP4, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t4<- sampTemp$t
        set.seed(i)
        shuffle <- sample(seq(nrow(samp)))
        samp <- samp[shuffle, ]
        elapsed <- as.numeric(difftime(Sys.time(), deb, units="secs"))
        save(samp, elapsed, file=title_file)
      }
    }
  }
  
}
rm(sampTemp)
gc()
```

## Case 2: sampling x that are not truly uniform but on a fine grid

```{r samplegeneration2, eval=FALSE}
i<- 1
for(i in seq(nRep)){
  for(nDiscret in c(51, 501)){
    for(nmax in round(10**seq(1, 5, 0.5))){
      title_file <- paste0("./samplingstudy/SLGPsSamp", i, 
                           "discret", nDiscret, "n", nmax,
                           "case2.Rdata")
      if(!file.exists(title_file)){
        deb<- Sys.time()
        cat(i, ".1\n")
        set.seed(i)
        x_candidates <- sample(seq(0, 1,, 1001), nmax, replace=TRUE)
        set.seed(i)
        t <- as.data.frame(table(x_candidates))
        sampTemp <- sampleSLGP(SLGPmodel=SLGP1, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp<- data.frame(x=sampTemp$x,
                          t1 = sampTemp$t)
        set.seed(i)
        cat(i, ".2\n")
        sampTemp <- sampleSLGP(SLGPmodel=SLGP2, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t2<- sampTemp$t
        set.seed(i)
        cat(i, ".3\n")
        
        sampTemp <- sampleSLGP(SLGPmodel=SLGP3, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t3<- sampTemp$t
        set.seed(i)
        cat(i, ".4\n")
        
        sampTemp <- sampleSLGP(SLGPmodel=SLGP4, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=as.numeric(as.character(t[, 1]))), 
                               n=t$Freq, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t4<- sampTemp$t
        set.seed(i)
        shuffle <- sample(seq(nrow(samp)))
        samp <- samp[shuffle, ]
        elapsed <- as.numeric(difftime(Sys.time(), deb, units="secs"))
        save(samp, elapsed, file=title_file)
      }
    }
  }
}
rm(sampTemp)
gc()
```


## Case 3: sampling x that are truly uniform

```{r samplegeneration3, eval=FALSE}
nmax <- 10000
i<- 1
nDiscret <- 51
for(i in seq(nRep)){
  for(nDiscret in c(51, 501)){
    for(nmax in round(10**seq(1, 4, 0.5))){
      
      title_file <- paste0("./samplingstudy/SLGPsSamp", i, 
                           "discret", nDiscret, "n", nmax,
                           "case3.Rdata")
      if(!file.exists(title_file)){
        deb<- Sys.time()
        cat(i, ".1\n")
        set.seed(i)
        x_candidates <- runif(nmax)
        set.seed(i)
        sampTemp <- sampleSLGP(SLGPmodel=SLGP1, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=x_candidates), 
                               n=1, 
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp<- data.frame(x=sampTemp$x,
                          t1 = sampTemp$t)
        set.seed(i)
        cat(i, ".2\n")
        sampTemp <- sampleSLGP(SLGPmodel=SLGP2, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=x_candidates), 
                               n=1,
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t2<- sampTemp$t
        set.seed(i)
        cat(i, ".3\n")
        
        sampTemp <- sampleSLGP(SLGPmodel=SLGP3, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=x_candidates), 
                               n=1,
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t3<- sampTemp$t
        set.seed(i)
        cat(i, ".4\n")
        
        sampTemp <- sampleSLGP(SLGPmodel=SLGP4, 
                               interpolateBasisFun="WNN",
                               newX=data.frame(x=x_candidates), 
                               n=1,
                               nIntegral=101,
                               nDiscret=nDiscret)
        samp$t4<- sampTemp$t
        set.seed(i)
        shuffle <- sample(seq(nrow(samp)))
        samp <- samp[shuffle, ]
        elapsed <- as.numeric(difftime(Sys.time(), deb, units="secs"))
        save(samp, elapsed, file=title_file)
      }
    }
  }
}
rm(sampTemp)
gc()


```


```{r, fig.cap = "Execution time on a laptop", fig.fullwidth=TRUE, fig.height=5, fig.width=8, fig.align='center' ,fig.pos="H", warning=FALSE, message=FALSE, echo=FALSE}
df_time <- data.frame(expand.grid(c(51, 501), seq(3), round(10**seq(1, 5, 0.5))))
colnames(df_time) <- c("nDiscret", "case", "n")
df_time$Discretisation <- paste0("CDF discretisation with ", df_time$nDiscret, " nodes.")
df_time$CaseText <- c("X sampled on a regular grid with 101 nodes.",
                      "X sampled on a regular grid with 1001 nodes.",
                      "X sampled on all of [0, 1].")[df_time$case]
df_time$time <- NA
for(i in seq(nrow(df_time))){
  file_title <- paste0("./samplingstudy/SLGPsSamp1discret", df_time$nDiscret[i], 
                       "n", df_time$n[i],
                       "case", df_time$case[i],".Rdata")
  if(file.exists(file_title)){
    load(file_title)
    df_time$time[i] <- elapsed
  }
}
df_time %>%
  dplyr::filter(!is.na(time))%>%
  ggplot(aes(x=n, y=time, col=CaseText))+
  geom_line()+
  geom_point()+
  facet_grid(.~factor(Discretisation, 
                      levels= paste0("CDF discretisation with ", 
                                     c(51, 501), " nodes.")))+
  theme_bw()+
  scale_x_log10()+
  labs(
    y = "Execution time [s]",
    x = "Sample size",
    col = "Sampling strategy"
  )


```
The blow-up in execution times for large samples is mostly due to the need to compute a normalisation term for all different values of $x$.

# Doing estimation 

```{r sampleload, eval=FALSE}

for(case in seq(3)){
  for(nDiscret in c(51, 501)){
    for(nFreq in c(25, 100, 500)){
      if(case %in% c(1, 2)){
        load(paste0("./samplingstudy/SLGPsSamp1discret", nDiscret, 
                    "n1e+05case", case, ".Rdata"))
        listN <- round(10**seq(1, 5, 1))
      }else{
        load(paste0("./samplingstudy/SLGPsSamp1discret", nDiscret, 
                    "n10000case", case, ".Rdata"))
        listN <- round(10**seq(1, 4, 1))
      }
      for(ref in seq(4)){
        modid <- ref
        for(nsamp in listN){
          title_file <- paste0("./samplingstudy/SLGP_ref1_mod1_freq", 
                               nFreq, "_n", nsamp, "_case", case,
                               "_nDiscret", nDiscret, "_ref", ref, ".RData")
          if(!file.exists(title_file)){
            cat(title_file, "\n")
            ind <- c(1:nsamp)
            if(ref==1){
              #exp
              current_samp <- data.frame(t=samp$t1[ind], x=samp$x[ind])
              refval <- pred1$pdf_1
              BasisFunParam <- SLGP1@BasisFunParam
            }
            if(ref==2){
              #Mat32
              current_samp <- data.frame(t=samp$t2[ind], x=samp$x[ind]) 
              refval <- pred2$pdf_1
              BasisFunParam <- SLGP2@BasisFunParam
            }
            if(ref==3){
              #Mat52            
              current_samp <- data.frame(t=samp$t3[ind], x=samp$x[ind]) 
              refval <- pred3$pdf_1
              BasisFunParam <- SLGP3@BasisFunParam
              
            }
            if(ref==4){
              #Gau
              current_samp <- data.frame(t=samp$t4[ind], x=samp$x[ind]) 
              refval <- pred4$pdf_1
              BasisFunParam <- SLGP4@BasisFunParam
            }
            BasisFunParam$freq <- BasisFunParam$freq[c(1:nFreq, 500+(1:nFreq)), ]
            BasisFunParam$coef <- rep(1/sqrt(nFreq), 2*nFreq)
            BasisFunParam$offset <- BasisFunParam$offset[c(1:nFreq, 500+(1:nFreq))]
            if(modid==4){
              mod <- slgp(t~x,
                          data=current_samp,
                          method ="MAP",
                          basisFunctionsUsed="RFF",
                          interpolateBasisFun="WNN",
                          hyperparams = list(lengthscale=c(0.1, 0.1), sigma2=0.5),
                          predictorsUpper= c(1),
                          predictorsLower= c(0),
                          responseRange= c(0, 1),
                          sigmaEstimationMethod = "none",
                          seed=1,
                          BasisFunParam=BasisFunParam,
                          opts_BasisFun = list(nFreq=nFreq,
                                               MatParam=Inf))
            }else{
              mod <- slgp(t~x,
                          data=current_samp,
                          method ="MAP",
                          basisFunctionsUsed="RFF",
                          interpolateBasisFun="WNN",
                          hyperparams = list(lengthscale=c(0.1, 0.1), sigma2=0.5),
                          predictorsUpper= c(1),
                          predictorsLower= c(0),
                          responseRange= c(0, 1),
                          sigmaEstimationMethod = "none",
                          seed=1,
                          BasisFunParam=BasisFunParam,
                          opts_BasisFun = list(nFreq=nFreq,
                                               MatParam=1/2+modid-1))
            }
            mod <- slgp(t~x,
                        data=current_samp,
                        method ="MAP",
                        basisFunctionsUsed="RFF",
                        interpolateBasisFun="WNN",
                        hyperparams = list(lengthscale=c(0.1, 0.1), sigma2=0.5),
                        predictorsUpper= c(1),
                        predictorsLower= c(0),
                        responseRange= c(0, 1),
                        sigmaEstimationMethod = "none",
                        seed=1,
                        BasisFunParam=BasisFunParam,
                        opts_BasisFun = list(nFreq=nFreq,
                                             MatParam=1/2))
            predMod <- predictSLGP_newNode(SLGPmodel = mod, newNodes = dfGrid2)
            predMod$ref <- refval
            dH <- (predMod %>%
                     group_by(x) %>%
                     summarise(value=mean((sqrt(pdf_1) - sqrt(ref))^2)/2) %>%
                     ungroup() %>%
                     summarise(value=sqrt(mean(value))))$value
            save(mod, dH, file=title_file)
          }
        }
      }
    }
  }
  
}
```


This code is to compile previously obtained results

```{r estimationcompile, eval=FALSE}
dfres <- data.frame()
listN <- round(10**seq(1, 5, 1))

for(case in seq(3)){
  for(nDiscret in c(51, 501)){
    for(nFreq in c(25, 100,500)){
      for(n in listN){
        title_file <- paste0("./samplingstudy/SLGP_ref1_mod1_freq", 
                             nFreq, "_n", n, "_case", case,
                             "_nDiscret", nDiscret, ".RData")
        if(file.exists(title_file)){
          load(file=title_file)
          tempdf <- data.frame(nFreq=nFreq,
                               case=c("X sampled on a regular grid with 101 nodes.",
                                      "X sampled on a regular grid with 1001 nodes.",
                                      "X sampled on all of [0, 1].")[case],
                               nDiscret=paste0("CDF discretisation with ", 
                                               nDiscret, " nodes."),
                               n=n,
                               dH=dH)
          dfres <- rbind(dfres, tempdf)
        }
      }
    }
  }
}

dfres$case <- factor(dfres$case,
                     levels=c("X sampled on a regular grid with 101 nodes.",
                              "X sampled on a regular grid with 1001 nodes.",
                              "X sampled on all of [0, 1]."))
dfres$nDiscret <- factor(dfres$nDiscret,
                         levels=paste0("CDF discretisation with ", 
                                       c(51, 501), " nodes."))
save(dfres, file="./samplingstudy/compiledRes.RData")
```


```{r finalplot, fig.fullwidth=TRUE, fig.height=8, fig.width=10, fig.align='center' ,fig.pos="H"}
load(file="./samplingstudy/compiledRes.RData")

dfres %>%
  filter(!is.na(dH))%>%
  ggplot()+
  geom_line(mapping=aes(x=n, y=dH, col=as.factor(2*nFreq), lty=nDiscret))+
  geom_hline(yintercept = 0, col="grey")+
  facet_grid(.~case)+
  scale_x_log10()+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        legend.position = "bottom",
        legend.direction = "horizontal"
  ) +
  labs(
    y = "Integrated Squared Hellinger distance",
    x = "Sample size",
    col = "Number of basis functions",
    fill = "Number of basis functions"
  )
```

```{r}
ggplot(sampTemp, aes(x=x, y=t) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )

ggplot(dfplot, aes(x=x, y=t, fill=pdf_1))+
  geom_raster()+
  facet_grid(.~kernel)+
  geom_point(samp, mapping=aes(x=x, y=t1), inherit.aes=FALSE, pch=".", col="white")
```

